---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# lizardHMM

<!-- badges: start -->
<!-- badges: end -->

The goal of lizardHMM is to fit lizard movement data with Hidden Markov Models.

## Installation

You can install the released version of lizardHMM from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("lizardHMM")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("simonecollier/lizardHMM")
```
## Example

Set up a basic HMM with data distributed according to normal state dependent 
distributions. This HMM will have 3 states, 2 variables, 1 subject, and 2 
covariates (included in the computation of the transition probability matrix).
We also define the design matrix which indicates the values of the covariates
at each point in time. Then we go about generating data from this HMM.

```{r hmm_setup}
library(lizardHMM)

hmm1 <- list(num_states = 3,
             num_variables = 2,
             num_subjects = 1,
             mu = list(matrix(c(12, 18, 22), ncol = 3, nrow = 1, byrow = TRUE), 
                       matrix(c(-12, -7, 0), ncol = 3, nrow = 1, byrow = TRUE)),
             sigma = list(matrix(c(3, 1, 1.5), ncol = 3, nrow = 1, byrow = TRUE), 
                          matrix(c(1, 3, 2), ncol = 3, nrow = 1, byrow = TRUE)),
             beta  = matrix(c(0.01, 0.02, 0.001,
                              0.01, 0.03, 0.004,
                              0.01, 0.01, 0.003,
                              0.01, 0.04, 0.002,
                              0.01, 0.01, 0.004,
                              0.01, 0.03, 0.001), ncol = 3, nrow = 6, byrow = TRUE),
             delta = list(c(0.3, 0.2, 0.5)))

num_sample <- 1000

design           <- list(matrix(0, nrow = 1000, ncol = 3))
design[[1]][, 1] <- 1 # First column is the intercept
design[[1]][, 2] <- sample(c(0, 1), size = 1000, 
                           prob = c(0.3, 0.7), replace = TRUE)
design[[1]][, 3] <- rnorm(1000, mean = 5, sd = 1)

sample <- norm_generate_sample(num_sample, hmm1, design)
x      <- sample$observ
```

We can plot the timeseries of the observations for each subect and variable 
along with the corresponding states.

```{r timeseries_plot, echo = TRUE}
timeseries_plot(x, sample$state, hmm1$num_subjects, hmm1$num_variables)
```

The histogram of the data for each subject and variable can be gnerated, 
overlayed with the state dependent normal distributions.

```{r norm_hist, echo = TRUE}
norm_hist(sample, hmm1$num_states, hmm1$num_variables, hmm1$num_subjects,
          hmm1, width = 1, x_step = 0.2)
```

Now we can try fitting the data we generated with a basic HMM with reasonable 
guesses for the initial parameters.

```{r hmm_fit, echo = TRUE, warning = FALSE, message = FALSE}
num_states = 3
num_variables = 2
num_subjects = 1
num_covariates = 2
mu0 <- list(matrix(c(11, 19, 23), ncol = 3, nrow = 1, byrow = TRUE),
            matrix(c(-10, -5, 1), ncol = 3, nrow = 1, byrow = TRUE))
sigma0 <- list(matrix(c(3, 3, 3), ncol = 3, nrow = 1, byrow = TRUE),
               matrix(c(3, 3, 3), ncol = 3, nrow = 1, byrow = TRUE))
beta0 <- matrix(c(-2, 0, 0,
                  -2, 0, 0,
                  -2, 0, 0,
                  -2, 0, 0,
                  -2, 0, 0,
                  -2, 0, 0), ncol = 3, nrow = 6, byrow = TRUE)
delta0 <- list(c(1/3, 1/3, 1/3))

hmm_fit <- norm_fit_hmm(x, design, num_states, num_variables, num_subjects,
                        num_covariates, mu0, sigma0, beta0, delta0,
                        iterlim = 200, hessian = TRUE)
```

We can find the confidence intervals for our parameter estimates.

```{r confidence_intervals, echo = TRUE}
conf_intervals <- norm_ci(hmm_fit, state_dep_dist_pooled = FALSE, 
                          n = 100, level= 0.975) 
conf_intervals
```

Using the viterbi algorithm we can decode the hidden states according to our 
fitted HMM and plot the resulting time series.

```{r viterbi, echo = TRUE}
viterbi <- norm_viterbi(x, hmm_fit)
timeseries_plot(x, viterbi, num_subjects, num_variables)
```

We can compute the confidence intervals of the state dependent normal 
distributions and compare them to the histogram of data.

```{r hist_ci, echo = TRUE}
norm_hist_ci(x, viterbi, num_states, num_subjects, num_variables, hmm_fit, 
             state_dep_dist_pooled = FALSE,
             width = 1, n = 100, level = 0.975, x_step = 0.2)
```





